# Differential expression analysis 

Having defined our statistical model, we will proceed with the differential expression analysis, using the `limma` package. The `voom` function transforms count data to log2-counts per million. From this, the mean-variance relationship is used to compute observation level weights. 

```{r}
library("limma")

vGene <- voom(dge, mod, plot = TRUE)
```

Our next step is calculating different statistics, to evaluate which genes are differentially expressed in different circumstances. For this, we will use the `eBayes` function. Looking back at what each coefficient represents:

```{r}
colnames(mod)
```

We will look at the genes differentially expressed with respect to coefficients 3 (Basal), 4 (Her2), 5 (LumA) and 6 (LumB), to look at the genes associated with the different subtypes. We can predict that these genes are among the ones used to distinguish the subtypes. 

```{r}
# Empirical Bayes Statistics for Differential Expression
eb_results <- eBayes(lmFit(vGene))
```

## Normal tissue  

```{r}
de_results_normal <- topTable(
    eb_results,
    coef = 1,
    number = nrow(rse_BRCA),
    sort.by = "none"
)

# Number of differentially expressed genes
table(de_results_normal$adj.P.Val < 0.05)
```
When compared to cancer, most genes are differentially expressed. 

## Basal subtype

```{r}
de_results_basal <- topTable(
    eb_results,
    coef = 3,
    number = nrow(rse_BRCA),
    sort.by = "none"
)

# Number of differentially expressed genes
table(de_results_basal$adj.P.Val < 0.05)
```
65% of genes are differentially expressed. 

```{r}
# Visualizing as volcano plot
volcanoplot(eb_results, coef = 3, highlight = 3, names = de_results_basal$gene_name)
```
```{r}
# Highlighted genes 
de_results_basal$gene_name[rank(de_results_basal$adj.P.Val) < 4]
```

## Her2 subtype

```{r}
de_results_her2 <- topTable(
    eb_results,
    coef = 4,
    number = nrow(rse_BRCA),
    sort.by = "none"
)

# Number of differentially expressed genes
table(de_results_her2$adj.P.Val < 0.05)
```
60% of genes are differentially expressed. 

```{r}
# Visualizing as volcano plot
volcanoplot(eb_results, coef = 4, highlight = 3, names = de_results_her2$gene_name)
```
```{r}
# Highlighted genes 
de_results_her2$gene_name[rank(de_results_her2$adj.P.Val) < 4]
```
## LumA subtype

```{r}
de_results_luma <- topTable(
    eb_results,
    coef = 5,
    number = nrow(rse_BRCA),
    sort.by = "none"
)

# Number of differentially expressed genes
table(de_results_luma$adj.P.Val < 0.05)
```
47% of genes are differentially expressed. 

```{r}
# Visualizing as volcano plot
volcanoplot(eb_results, coef = 5, highlight = 3, names = de_results_luma$gene_name)
```

```{r}
# Highlighted genes 
de_results_luma$gene_name[rank(de_results_luma$adj.P.Val) < 4]
```

## LumB subtype

```{r}
de_results_lumb <- topTable(
    eb_results,
    coef = 6,
    number = nrow(rse_BRCA),
    sort.by = "none"
)

# Number of differentially expressed genes
table(de_results_lumb$adj.P.Val < 0.05)
```
63% of genes are differentially expressed. 

```{r}
# Visualizing as volcano plot
volcanoplot(eb_results, coef = 6, highlight = 3, names = de_results_lumb$gene_name)
```


```{r}
# Highlighted genes 
de_results_lumb$gene_name[rank(de_results_lumb$adj.P.Val) < 4]
```